---
title: "Homework 01"
---

## TidyTuesday Part

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge.
Develop a research question, then answer it through a short data story mainly composed of [effective visualizations]({{< var effective-viz-url >}}).
Provide sufficient background for readers to grasp your narrative.
Refer to the Homework Instructions ⭐ page on the course website for details on how to submit this part.

> Cappuccino prices differ from cafe to cafe, but prices may also differ across countries.
> The TidyTuesday dataset contains information about individual cafes, including the price of a small cappuccino in both the original currency and British pounds.
> My research question is simple; "do cappuccino prices vary substantially between countries?"

## Methodology

> Lets load the data;

```{r}
library(tidyverse)

cafe <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-09-08/cafe.csv"
)

cafe$country <- ifelse(cafe$country == "Turkey", "Turkiye", cafe$country)
```

> We need to clean the data because the dataset has an invisible character in the country names for **New Zealand and South Africa**.
> The TidyTuesday documentation specifically warns about this because it can cause joins to fail.

```{r}
#The \u00a0 is a non-breaking space that occurs in the country names for New Zealand and South Africa.

cafe <- cafe |>
  mutate(country = gsub("\u00a0", " ", country))

```

> Looking at our dataset...

```{r}
glimpse(cafe)
```

> Lets look at how many cafes do we have from each country?

```{r}
count(cafe, country)
```

> What does the cappuccino price variable look like?

```{r}
summary(cafe$price_gbp)
```

## Visualizations

1.  Calculate the average price for each country

```{r}
country_prices <- cafe |>
  group_by(country) |>
  summarise(
    mean_price = mean(price_gbp, na.rm = TRUE)
  )

country_prices

```

### Graph 1- The overall picture

> What does the overall distribution of country-level cappuccino prices look like?

```{r}
ggplot(country_prices,
       aes(x = mean_price)) +
  geom_histogram(binwidth = 0.5, fill= "pink") +
  labs(
    title = "Cappuccino Prices Differ Across Countries",
    subtitle = "Distribution of average cappuccino prices by country",
    x = "Average cappuccino price (£)",
    y = "Number of countries"
  ) +
  theme_minimal()

```

> The above graph indicates that the cappuccino prices vary, with a range of \~0.5£ to \~5.75£ with one country having average cappuccino price above 6 pounds.
> It is fairly distributed, most countries having average price of around 2-3£.
> Although this graph gives us an overview, we will still need to investigate further.

### Graph II: Show the extremes

```{r}
cheapest <- country_prices |>
  slice_min(mean_price, n = 10)


most_expensive <- country_prices |>
  slice_max(mean_price, n = 10)


price_extremes <- bind_rows(
  cheapest,
  most_expensive
)
# Note to self: bind_rows will put these datasets on top of eachother, reorder will not put the countries in alphabetical order!! geom_col() is better when you know the values calculated in which I did above.

ggplot(
  price_extremes,
  aes(x = mean_price, y = reorder(country, mean_price)
  )
) +
  geom_col(fill= "orange") +
  labs(
    title = "The Cheapest and Most Expensive Cappuccinos",
    subtitle = "Countries with the 10 lowest and 10 highest average prices",
    x = "Average cappuccino price (£)",
    y = NULL
  ) +
  theme_minimal()
```

> The countries at the extremes show that the difference in average cappuccino prices can be substantial.
> The lowest-priced countries have average prices around £1–£2, while the highest-priced countries have averages above £4, with some exceeding £5.

### Graph III: Is the variation only between countries, or do prices also vary substantially **within** countries?

```{r}
country_counts <- cafe |>
  count(country, sort = TRUE)

country_counts

countries_selected <- c(
  "Denmark",
  "Belgium",
  "Brazil",
  "Turkiye",
  "India", 
  "Croatia"
)

cafe_selected <- cafe |>
  filter(country %in% countries_selected)

ggplot(
  cafe_selected,
  aes(
    x = price_gbp,
    y = reorder(country, price_gbp, FUN = median)
  )
) +
  geom_boxplot(
    outlier.shape = NA
  ) +
  geom_jitter(
    height = 0.1,
    alpha = 0.4,
    color = "black"
  ) +
  labs(
    title = "Cappuccino Prices Vary Within Countries",
    subtitle = "Denmark has the highest typical prices and greatest variation among the countries shown",
    x = "Cappuccino price (£)",
    y = NULL,
    caption = "Countries were selected based on café sample size. Source: Cappuccino Index dataset, TidyTuesday. Created by Riley Karatas | September 26, 2026."
  ) +
  theme_minimal()
```
> Although the country averages reveal differences between countries, individual cafe prices show that prices can also vary within a country.
> For example, the sampled cafes in **Denmark** show a considerably wider range of prices than those in **Croatia** and **Brazil**.
> Denmark also contains several higher-priced observations that extend well beyond the main cluster of cafes.
> This demonstrates why looking only at a country's average can be misleading: an average summarizes the data, but it does not show the full range of prices found among individual cafes.
