---
title: "Homework 01"
---

## TidyTuesday Part

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge.
Develop a research question, then answer it through a short data story mainly composed of [effective visualizations]({{< var effective-viz-url >}}).
Provide sufficient background for readers to grasp your narrative.
Refer to the Homework Instructions ⭐ page on the course website for details on how to submit this part.

> Cappuccino prices differ from cafe to cafe, but prices may also differ across countries.
> The TidyTuesday dataset contains information about individual cafes, including the price of a small cappuccino in both the original currency and British pounds.
> My research question is simple; "do cappuccino prices vary substantially between countries?"

## Methodology

> Lets load the data;

```{r}
library(tidyverse)

cafe <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-09-08/cafe.csv"
)
```

> We need to clean the data because the dataset has an invisible character in the country names for **New Zealand and South Africa**.
> The TidyTuesday documentation specifically warns about this because it can cause joins to fail.

```{r}
#The \u00a0 is a non-breaking space that occurs in the country names for New Zealand and South Africa.

cafe <- cafe |>
  mutate(country = gsub("\u00a0", " ", country))

```

> Looking at our dataset...

```{r}
glimpse(cafe)
```

> Lets look at how many cafes do we have from each country?

```{r}
count(cafe, country)
```

> What does the cappuccino price variable look like?

```{r}
summary(cafe$price_gbp)
```

## Visualizations

```{r}
ggplot(cappuccino_index, 
       aes(x = reorder(country, index), 
           y = index)) +
  geom_col() + 
  coord_flip() + 
  labs(
    title = "How long does a coppucino cost a barista?",
    x = NULL, 
    y= "Minutes of work"
  )
```

```{r}
ggplot(cappuccino_index,
       aes(x = n, y = index)) +
  geom_point() +
  scale_x_log10() +
  labs(
    title = "How Much Evidence Supports Each Country's Index?",
    x = "Number of cafes sampled",
    y = "Cappuccino Index (minutes)"
  ) +
  theme_minimal()
```

```{r}
ggplot(cappuccino_index,
       aes(x = n, y = index)) +
  geom_point() +
  scale_x_log10() +
  labs(
    title = "Cappuccino Indexes Are Based on Different Sample Sizes",
    x = "Number of cafes sampled (log scale)",
    y = "Minutes of barista work"
  ) +
  theme_minimal()
```

```{r}
ggplot(cafe,
       aes(x = country,
           y = price_gbp)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Cappuccino Prices Vary Within Countries",
    x = NULL,
    y = "Cappuccino price (£)"
  ) +
  theme_minimal()
```
