---
title: "Personal Visualization Journey"
---

# A Year in My Soundtrack

> How did my music obsessions change throughout the year?

> Let's load the data and read in the json files I received from Spotify website.

```{r}
library(jsonlite)
library(dplyr)
library(purrr)

files <- list.files(
  "data/Spotify_Account_Data",
  pattern = "StreamingHistory_music_.*\\.json$",
  full.names = TRUE
)

files

spotify <- map_dfr(files, fromJSON)

glimpse(spotify)
```

> Now we need to clean the data in a way we can understand and analyze better!

```{r}
spotify_clean <- spotify %>%
  select(
    endTime,
    artistName,
    trackName,
    msPlayed
  ) %>%
  mutate(
    datetime = as.POSIXct(endTime, format = "%Y-%m-%d %H:%M", tz = "UTC"),
    date = as.Date(datetime),
    hour = as.integer(format(datetime, "%H")),
    day = format(datetime, "%A"),
    month = format(datetime, "%Y-%m"),
    minutesPlayed = msPlayed / 60000
  )

glimpse(spotify_clean)
```

```{r}
range(spotify_clean$date, na.rm = TRUE)
```

> We have almost exactly a year worth of listening data!

```{r}
sum(spotify_clean$minutesPlayed, na.rm = TRUE)
```

```{r}
spotify_clean %>%
  count(artistName, sort = TRUE) %>%
  head(20)
```

> And this is my 20 most frequent played artists!

```{r}
artist_minutes <- spotify_clean %>%
  filter(artistName != "Unknown Artist") %>%
  group_by(artistName) %>%
  summarise(
    minutes = sum(minutesPlayed, na.rm = TRUE),
    plays = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(minutes))

head(artist_minutes, 20)
```

> This gives us my actual top artists by listening time.

> Now let's calculate artist and the month

```{r}
monthly_artist <- spotify_clean %>%
  filter(artistName != "Unknown Artist") %>%
  group_by(month, artistName) %>%
  summarise(
    minutes = sum(minutesPlayed, na.rm = TRUE),
    .groups = "drop"
  )

top_artists <- artist_minutes %>%
  slice_head(n = 12) %>%
  pull(artistName)

monthly_top <- monthly_artist %>%
  filter(artistName %in% top_artists)

monthly_top
```

> Let's start visualizing

```{r}
library(ggplot2)
library(tidyr)

monthly_top <- monthly_artist %>%
  filter(artistName %in% top_artists) %>%
  complete(
    month,
    artistName,
    fill = list(minutes = 0)
  )

ggplot(
  monthly_top,
  aes(
    x = month,
    y = artistName,
    fill = minutes
  )
) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "C",
    name = "Minutes listened"
  ) +
  labs(
    title = "A Year in My Soundtrack",
    subtitle = "Amy Winehouse is my top listened artist this year",
    x = NULL,
    y = NULL, 
    caption = "*My top 12 frequently listened artists and their change against months. Data: Spotify User Data, Spotify, Week of October 5th, 2026.\nCreated by Riley Karatas | Creation date: October 5th, 2026."
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

```

> The conlcusion is that I listen to a lot of Amy Winehouse depending on the month. It seems that my taste in music is quite international with Russian, Turkish, French Afro-beat and English songs.
